---
title: "2021 Lab Journal"
author: "Kelly Mistry"
date: "1/26/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Applied Time Series Analysis course materials

Continuing to work through Applied Time Series Analysis course materials. 

#### Section Five

Box-Jenkins method steps:
* Model form selection, which entails evaluating stationarity (and if stationarity needs to be adjusted, selecting differencing level), selecting the AR level (p) and selecting the MA level (q)
* Parameter estimation
* Model checking

Stationarity tests: Dickey-Fuller (for lag of 0), Augmented Dickey-Fuller (for lags greater than 0) and KPSS

```{r ATSA Section 5, eval = FALSE, echo = FALSE}

```

#### Section Six

```{r ATSA Section 6, eval = FALSE, echo = FALSE}
```

#### Section Seven

```{r ATSA Section 7, eval = FALSE, echo = FALSE}
```

#### Section Eight

```{r ATSA Section 8, eval = FALSE, echo = FALSE}
```

#### Section Nine

```{r ATSA Section 9, eval = FALSE, echo = FALSE}
```

#### Section Ten

```{r ATSA Section 10, eval = FALSE, echo = FALSE}
```

#### Section Eleven

```{r ATSA Section 11, eval = FALSE, echo = FALSE}
```


### Indigenous Statistics book notes continued




