---
title: "2021 Lab Journal"
author: "Kelly Mistry"
date: "1/26/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Applied Time Series Analysis course materials

Continuing to work through Applied Time Series Analysis course materials.

#### Section Five

Box-Jenkins method steps: \* Model form selection, which entails evaluating stationarity (and if stationarity needs to be adjusted, selecting differencing level), selecting the AR level (p) and selecting the MA level (q) \* Parameter estimation \* Model checking

Stationarity tests: Dickey-Fuller (for lag of 0), Augmented Dickey-Fuller (for lags greater than 0) and KPSS

```{r ATSA Section 5, eval = FALSE, echo = FALSE}

```

#### Section Six

```{r ATSA Section 6, eval = FALSE, echo = FALSE}
```

#### Section Seven

```{r ATSA Section 7, eval = FALSE, echo = FALSE}
```

#### Section Eight

```{r ATSA Section 8, eval = FALSE, echo = FALSE}
```

#### Section Nine

```{r ATSA Section 9, eval = FALSE, echo = FALSE}
```

#### Section Ten

```{r ATSA Section 10, eval = FALSE, echo = FALSE}
```

#### Section Eleven

```{r ATSA Section 11, eval = FALSE, echo = FALSE}
```

### Indigenous Statistics book notes continued

### Applied Time Series Analysis Winter 2021 Class Notes

-   Don't include co-linear covariates for MARSS models (multivariate Autoregressiv State Space - same as Vector Autoregressive)
-   

## VAST project model runs, notes & plans

#### 5/18/2021

-   Set up code to create input objects (with POP only for now):

```{r VAST set up}
## Load packages and set working directory
library(TMB)               
library(VAST)
library(here)
library(dplyr)

# current VAST version this code was written for: 3.5.2
# current TMB version this was written for: 1.7.18

## Set species 
Species <- c("Sebastes alutus", "Sebastes polyspinis", "Pleurogrammus monopterygius")[1] 

## Set up folder to store results
results_folder <- paste0(here("results"), "/", Species[1])
dir.create(results_folder)

## Load raw survey data
raw_haul <- read.csv(here("data", "haul.csv"))

## Define stock names & NOAA ID numbers
stock_names <- c("Pacific Ocean Perch" = "Sebastes alutus", "Northern Rockfish" = "Sebastes polyspinis", "Atka Mackerel" = "Pleurogrammus monopterygius")
my_stock_ids <- c("30060", "30420", "21921")
names(my_stock_ids) <- stock_names

# Pre-processed data provided by Cecilia \
CPUE_data <- read.csv(here("data/CPUE.csv"))
BM_Stratum_data <- read.csv(here("data/BIOMASS_STRATUM.csv"))

# Separating out one species, POP is the first stock
CPUE_data <- CPUE_data[CPUE_data$SPECIES_CODE == my_stock_ids[1],]

# for POP species, start at 1990, for northern rockfish start at 1984:
CPUE_data <- CPUE_data[CPUE_data$YEAR >= 1990, ] # for POP
# CPUE_data <- CPUE_data[CPUE_data$YEAR >= 1984, ] # for northern rockfish

# Adding in start longitude and latitude from the raw_haul data frame into the CPUE_data data frame
CPUE_data <- merge(CPUE_data, raw_haul[, c("HAULJOIN", "START_LATITUDE", "START_LONGITUDE")], by = "HAULJOIN")

## Separating out the data for VAST input object, which is named Data_Geostat, and has to have these columns
Data_Geostat <- CPUE_data[, c("WEIGHT", "YEAR", "EFFORT", "START_LATITUDE", "START_LONGITUDE")]

# Renaming column names to the ones required by VAST
colnames(Data_Geostat) <- c("Catch_KG", "Year", "AreaSwept_km2", "Lat", "Lon")
```

-   Two versions of the beta term (temporal coefficient) in RhoConfig setting

    -   beta_1 and beta_2 = 0, indicating that the model should treat time as a fixed intercept

    -   beta_1 and beta_2 = 4, indicating that the model should treat the temporal coefficient as an autoregressive 1 process (1 time step)

-   Two distributions for the error distribution that need to be tried out and tested to see which fits the data better, the gamma distribution and the Tweedie distribution; this is controlled by an argument in the make_settings function

    -   ObsModel = c(2, 1) is the gamma distribution for positive catch, delta model for encounter probability

    -   ObsModel = c(10, 2) is tweedie for positive catch and tweedie link for encounter

-   When I run the model, some of the outputs go into the folder that I name, but all of the plots go into the folder that has the script in it. Look deeper in the code to figure out why this is happening and see if I can do a work around and/or reach out to Cecilia and Jim to let them know that this is happening so they can fix it in the next update. Maybe create a github issue about it?

-   Update VAST to most recent edition before I do anything else. My current code (with betas = 4, and set to gamma dist) is:

```{r VAST code 5.18.21}
# Defining the 3 sub-regions
strata.limits <- data.frame(STRATA = as.factor(c("Western", "Central", "Eastern")),
                            west_border = c(-Inf, -159, -147),
                            east_border = c(-159, -147, Inf))

## Spatial & spatiotemp random effects settings - ***I think these stay the same…
FieldConfig <- matrix( c("IID","IID","IID",
                         "IID","IID","IID"), #"IID" = independent 
                       ncol=2, nrow=3, 
                       dimnames=list(c("Omega","Epsilon","Beta"), 
                                     c("Component_1","Component_2")) ) 

#omega = spatial factor for encounter[1] & positive catch[2]
#epsilon = spatiotemp factor for encounter[1] & positive catch[2]

#beta = number of factors for intercepts (relevant for multivariate models - factor analysis for years)

## Temporal settings  
RhoConfig_betas <- c(4, 4) #run two versions: 0,0 for intercept and 4,4 for AR1 process

RhoConfig  <- c("Beta1" = RhoConfig_betas[1], "Beta2" = RhoConfig_betas[2], "Epsilon1" = 0, "Epsilon2" = 0)

## beta = autocorr for ecounter [1] & positive catch [2], zero is fixed effect intercept
## epsilon = spatiotemp variation for encounter[1] & positive catch[2]

settings = make_settings( Version = "VAST_v8_2_0", #.cpp version, not software #e.g., "VAST_v12_0_0"
                          n_x = 500, #knots aka spatial resolution of our estimates
                          Region = "User", 
                          purpose = "index2", #changes default settings
                          ObsModel= c(2,1), #c(1,1) #c(10,2); (2,1) is gamma for positive catch, delta model for encounter, and (10, 2) is tweedie for positive catch and tweedie link for encounter
                          ## everything after this is default if you use purpose = "index2"##
                          FieldConfig = FieldConfig, #spatial & spatiotemp random effects 
                          RhoConfig = RhoConfig, #temporal settings; default is all 0s, but if I specify this it will be changed here
                          strata.limits = strata.limits, #define area that you're producing index for
                          "knot_method" = "grid", #knots in proportion to spatial domain #other option is knot_method="samples"
                          fine_scale = TRUE, #changes the type of interpolation in your extrapolation area
                          bias.correct = TRUE, #corrects the index for nonlinear transformation; I want this for the final version, but I can turn it off while I'm messing with the model so it will run faster
                          use_anisotropy = TRUE) ##correlations decline depend on direction if this argument is TRUE

## Import extrapolation grid, these are available on Jason's Google drive: VASTGAP\Extrapolation Grids
GOAgrid <- read.csv(file= paste0(here("/data"),"/Extrapolation_Grids/GOAThorsonGrid_Less700m.csv"))

input_grid <- cbind(Lat = GOAgrid$Lat,
                    Lon = GOAgrid$Lon,
                    Area_km2 = GOAgrid$Shape_Area/1000000)  # Extrapolation grid area is in 
                                                            # m^2 & is converted to km^2

gc() #garbage collector

## Run model

fit <- fit_model( "settings"= settings, #all of the settings we set up above
                 "Lat_i"= Data_Geostat[,'Lat'], #latitude of observation
                 "Lon_i"= Data_Geostat[,'Lon'],  #longitude of observation
                 "t_i"= Data_Geostat[,'Year'], #time for each observation
                 "c_i"= rep(0,nrow(Data_Geostat)), #categories for multivariate analyses; don't actually use this, could comment it out if it doesn't have a fit
                 "b_i"= Data_Geostat[,'Catch_KG'], #in kg, raw catch or in CPUE per tow
                 "a_i"= Data_Geostat[,'AreaSwept_km2'], #sampled area for eahc observation
#                 "v_i"= Data_Geostat[,'Vessel'], #ok to leave in because it's all "missing" in data, so no vessel effects
                 "input_grid"= input_grid, #only needed if you have a user input extrapolation grid
                 "optimize_args" =list("lower"=-Inf,"upper"=Inf), #TMB argument (?fit_tmb)
                 "working_dir" = results_folder)

## Plot results
plot( fit )

## ## 
## save the VAST model
saveRDS(fit,file =  paste0(results_folder, "/", Species,"VASTfit.RDS"))
```

#### 5/20/2021

I'm updating VAST to the most recent version, which is: 3.7.1

-   Tried to run the gamma version (which worked with the 3.5.2 version) and got this error message: Error in (function (b_i, a_i, t_i, c_iz = rep(0, length(b_i)), e_i = c_iz[, : `X1_formula` and `X2_formula`; to use these please use `Version='VAST_v12_0_0'` or higher

    -   Figured out that this was referring to an argument in the make_settings() function, so I made it Version = "VAST_v12_0\_0" (it was previously Version = "VAST_v8_2\_0", in the run above)

-   The second attempt to run the gamma version has been going quite a bit longer than when I previously ran it (about 30 - 40 minutes) - I think I'll wait another hour (its been about an hour now) before stopping it. There haven't been any error messages, although this warning did appear a few lines ago: Note that \`getReportCovariance=FALSE\` causes an error in \`TMB::sdreport\` when no ADREPORTed variables are present

    -   The final line that is currently showing is: Bias correcting 90 derived quantities

    -   It finished, and produced the appropriate objects and plots, so I think it worked fine?

-   First attempt to run with the Tweedie distribution for errors (with ObsModel= c(10,2) in the make_settings function); seems to be going well so far.

    -   As the coefficients are scrolling by (presumably working on convergence?), and occasional message that says NA/NaN function evaluation appears, which didn't happen during any of the gamma runs. Check to see if this is anything to worry about with Cecilia.

    -   It is still running and seems okay, but it is definitely taking longer than the gamma versions - I've added code to record the run time for this script in the future, so I can actually know how long each script takes rather than having just a vague idea.

    -   This message also appears periodically, although it seems to be a general warning rather than a reaction to something in my code: Note that \`getReportCovariance=FALSE\` causes an error in \`TMB::sdreport\` when no ADREPORTed variables are present

    -   The model finished running, and produced all of the result objects and plots that I would expect, although they saved in several seemingly random locations, and some of the plots were named with "Plots" and then the plot title, not sure where that is coming from.

#### 5/21/2021

-   Tried running the model with tweedie distribution and betas = 4 (ObsModel = c(10,2) in make_settings function and "Beta1" = 4 and "Beta2" = 4 in RhoConfig object); it threw this error after starting to run the model and stopped:

```{r 5.21.21 VAST run message}
### Making TMB object
Note: Using Makevars in /Users/kellymistry/.R/Makevars 
make: Nothing to be done for `all'.
Constructing atomic tweedie_logW
Constructing atomic tweedie_logW
List of estimated fixed and random effects:

### Testing model at initial values
Error in if (any(Gradient0 == 0)) { : 
  missing value where TRUE/FALSE needed
```

-   Ask Cecilia what this means, and how to fix it.
